---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VIVECampbell:  Functions for Campbell reviews in VIVE

<!-- badges: start -->
<!-- badges: end -->

The package provides functions used in Campbell review conducted by [VIVE - The Danish Center for Social Science Research](https://www.vive.dk/en/). See also the VIVE Campbell [homepage](https://www.vive.dk/da/centre-og-netvaerk/campbell/).

## Installation

You can install the development version of VIVECampbell like so:

``` {r, eval = FALSE}

# install.packages("devtools")
devtools::install_github("MikkelVembye/VIVECampbell")

```

## Example

Calcualting degrees of freedom for studies with clustering in one treatment group only (Hedges & Citkowicz, 2015)

```{r}
library(VIVECampbell)
## basic example code

h_eq7_hedges2105 <- 
  df_h_1armcluster(N_total = 100, ICC = 0.1, N_grp = 60, avg_grp_size = 5)

h_eq7_hedges2105
```


# References

Hedges, L. V., & Citkowicz, M (2015). Estimating effect size when there is clustering in one treatment groups. *Behavior Research Methods*, 47(4), 1295-1308. <https://doi.org/10.3758/s13428-014-0538-z>
