---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VIVECampbell: Functions for Campbell reviews in VIVE

<!-- badges: start -->
<!-- badges: end -->

The package provides functions used in Campbell reviews and meta-analyses conducted by [VIVE - The Danish Center for Social Science Research](https://www.vive.dk/en/). See also the VIVE Campbell [homepage](https://www.vive.dk/da/centre-og-netvaerk/campbell/).

## Installation

You can install the development version of VIVECampbell like so:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("MikkelVembye/VIVECampbell")
```

```{r, message=FALSE}
# Load necessary packages
library(VIVECampbell)
library(dplyr)
```

## Example - Degrees of freedom calculation for cluster-designed studies

Calcualting degrees of freedom for studies with clustering in one treatment group only (Hedges & Citkowicz, 2015)

```{r}
## Degrees of freedom calculation for cluster bias correction when there is clustering in one treatment group only

h_eq7_hedges2015 <- 
  df_h_1armcluster(N_total = 100, ICC = 0.1, N_grp = 60, avg_grp_size = 5)

h_eq7_hedges2015
```

## Example - Data included in package

Data from the 'Targeted school-based interventions for improving reading and mathematics for students with or at risk of academic difficulties in Grades K-6: A systematic review' (Dietrichson et al., 2021)

```{r}

# Data from a Campbell Systematic Review conducted by Dietrichson et al. (2021)
glimpse(Dietrichson2021_data[1:20])

# Find documentation behind the data set by removing the below #
#?Dietrichson2021_data

```


# References

Dietrichson, J., Filges, T., Seerup, J. K., Klokker, R. H., Viinholt, B. C. A., Bøg, M., & Eiberg, M. (2021). Targeted school-based interventions for improving reading and mathematics for students with or at risk of academic difficulties in Grades K-6: A systematic review. *Campbell Systematic Reviews*, 17(2), e1152. <https://doi.org/10.1002/cl2.1152>

Hedges, L. V., & Citkowicz, M (2015). Estimating effect size when there is clustering in one treatment groups. *Behavior Research Methods*, 47(4), 1295-1308. <https://doi.org/10.3758/s13428-014-0538-z>
