---
title: "Entering data, calculating effect sizes, and cluster bias correction when there is clustering in one group only"
date: "Last modified `r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    number_sections: true
    toc: true
vignette: >
  %\VignetteIndexEntry{Entering data, calculating effect sizes, and cluster bias correction when there is clustering in one group only}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{dplyr}
  %\VignetteDepends{tibble}
  %\VignetteDepends{purrr}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(pillar.sigfig = 4) # ensure tibble include 4 digits
options(tibble.width = Inf)
options(dplyr.print_min = 310)
options(scipen = 10)
options(dplyr.summarise.inform = FALSE)
```

## Introduction

This vignette/article provides examples of how to enter data from rather complex result tables and how to calculate effect size and conduct cluster bias correction for such data, using functions from the `VIVECampbell` R package. 

## Loading necessary packages

```{r setup, message=FALSE}
library(VIVECampbell)
library(purrr)
library(dplyr)
```

## Enter data from Fisher & Bentley (1996)

The practice, we often use when we enter data for studies is to name the data object using the first authors surname and the year of the publication, i.e., an object with the following structure: `AuthorYear`. For the present study used in this vignette this guideline means that we will coin the data object `Fisher1996`. One important part of calculating effect size is to make clear documentation of how and from where we obtained data we use for effect size calculation. At a minimum you much write some like "Find data in Table 4 (p. 1248)" before the codes/chunk in which construct the result data. In some cases when studies provide enough information to calculation the pre-post test correlation, $\rho$, the first calculation can be done when entering the data. When studies, such as Fisher (insert reference), report raw pre and post test means and standard deviations (SDs) plus the mean difference and its standard error, we can use the correlation formulas given in [Chapter 6 of the Cochrane Handbook](https://training.cochrane.org/handbook/current/chapter-06). For the treatment group, we can calculated the pre-post test correlation, $\rho$ via

$$ \rho_T = \frac{SD^2_{T,pre} + SD^2_{T,post} - SD^2_{T,diff}}{2 \times SD_{T,pre} \times SD_{T,post}} $$
and for the control group

$$ \rho_C = \frac{SD^2_{C,pre} + SD^2_{C,post} - SD^2_{C,diff}}{2 \times SD_{C,pre} \times SD_{C,post}} $$
We will later on pool these correlation estimates. To do so - INTRODUCE log transformation

```{r}

Fisher1996 <- tibble::tibble(
  
  Setting = rep(c("Inpatient", "Outpatient"), each = 3, 2),
  
  Outcome = rep(c("Alcohol use", "Drug use"), each = 6),
  
  treatment = rep(rep(c("Disease", "Cognitive", "TAU"), 2), 2),
  
    # N is imputed/guested for illustrative purposes only, because we only know the total sample 
    # of the treatment groups (N = 19). We here assume that the control group have similar size as
    # the treatment group. 
    N = 9, 
    
    m_pre = c(
      .469, .441, .349, # Alcohol use - inpatient
      .725, .598, .682, # Alcohol use - outpatient
      
      .107, .116, .117, # Drug use - inpatient 
      .219, .200, .322  # Drug use - outpatient
    ),
    
    sd_pre = c(
      .12, .13, .22,
      .11, .16, .17,
      
      .09, .10, .12,
      .11, .12, .05
      
    ),
    
    m_post = c(
      .070, .018, .141,
      .521, .152, .492,
      
      .001, .008, .087,
      .167, .044, .216
      
    ),
    
    sd_post = c(
      .11, .05, .21,
      .11, .15, .27,
      
      .01, .02, .12,
      .10, .05, .15
      
      
      
    ),
    
    m_diff = c(
      .399, .423, .208,
      .204, .446, .190,
      
      .106, .108, .030,
      .052, .196, .106
      
    ),
    
    sd_diff = c(
      .02, .09, .04,
      .01, .02, .11,
      
      .09, .09, .02, 
      .02, .08, .11
      
    ),
  
    # Here we test if the mean differences reported in the study are in line
    # with the differences between the reported pre and post means. It seems to
    # be the case that they make a reporting error for the cognitive outpatient 
    # group mean difference on drug use. Wrtie about the error can be on many levels
    mean_diff_test = m_pre - m_post,
    
    # From the Cochrane handbook (Higgins & Thomas, 2019, p. 166)
    r = (sd_pre^2 + sd_post^2-sd_diff^2)/(2 * sd_pre * sd_post ),
    
    # Obtained from Wilson (2016). Converting sd_diff to raw sd
    sdg = sd_diff/sqrt(2*(1-r)),
    
    z = 0.5 * log( (1+r)/(1-r) ),
    v = 1/(N-3)
    
    
); Fisher1996

```

## Calculating Cohens' *d* and Hedges' *g*

Add text

```{r}

treat_label_fisher <- unique(Fisher1996$treatment)[1:2]


fisher_function <- function(label){
  
  Fisher1996 |> 
    dplyr::filter(treatment != label) |> 
    group_by(Setting, Outcome) |> 
    summarise(
      study = "Fisher1996",
      treatment = treatment[1],
      es_method = "diff-in-diffs",
      
      M = sum(v*z)/sum(v),
      ppcor = (exp(2*M)-1)/(exp(2*M)+1),
      
      Nt = N[1],
      Nc = N[2],
      N_tot = Nt + Nc,
      
      df_ind = N_tot,
      
      s_pool = sqrt(sum((N - 1) * sd_post^2) / df_ind),
      
      diff_t = m_post[1] - m_pre[1],
      diff_c = m_post[2] - m_pre[2],
      
      d = (diff_t - diff_c) / s_pool,
      vd = sum(1/N) * (2*(1-ppcor)) + d^2/(2*df_ind),
      se = sqrt(vd)
      
      
    ) |> 
    ungroup() |> 
    relocate(study)
  
}

Fisher1996_est <- 
  map_dfr(treat_label_fisher, ~ fisher_function(label = .x))

Fisher1996_est
```

## Cluster bias correction when there is clustering in the treatment group only

Add text

```{r}

# [INSERT EXAMPLE WHEN FUNCTIONS ARE ADDED TO THE PACKAGE]

```

## Reference 

Fisher, M. S., & Bentley, K. J. (1996). Two group therapy models for clients with a dual diagnosis of substance abuse and personality disorder. *Psychiatric services*, 47(6) <https://doi.org/10.1176/ps.47.11.1244>
